---
title: "p8105_hw3_hz3023"
author: "Huiyi Zhu"
date: "2025-10-12"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare
```{r}
library(ggridges)
library(tidyverse)
```

## Problem 1
```{r cars}

library(p8105.datasets)
data("instacart")
view("instacart") 
```
In this dataset, there are `r nrow(instacart)` observations and `r ncol(instacart)` variables. Key variable include the order id, product id, reordered times, user id and order number and others.

```{r}
instacart |>
  group_by(aisle) |>
  summarize(
    items = n()) |>
  mutate(
    aisle_ranking = min_rank(items)) |> 
  arrange(aisle_ranking)
```
There are 143 aisles in the data set. The most items ordered from fresh vegetables aisles.

Below is a bar polt show the number orderd in each aisle.
```{r}
instacart |>
  group_by(aisle) |>
  summarize(items = n()) |>
  filter(items > 10000) |>
ggplot(aes(x = aisle, y = items)) +
  geom_col()+
    coord_flip()
ggsave("p8105_hw1p1_barpolt.pdf")
```

```{r}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle, product_name) |>
  summarize(orders = n()) |> 
  arrange(aisle, desc(orders)) |>
  group_by(aisle) |>
  slice_head(n = 3) |>
  knitr::kable()
```

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) |>
  knitr::kable(digits = 1)
```

## Problem 2
```{r}
nyc_zillow_df = 
  read_csv("./zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv") |> 
  pivot_longer(
    -(RegionID:CountyName),
    names_to = "month",
    values_to = "price") |> 
  janitor::clean_names() |> 
  rename(zip_code = region_name) |> 
  mutate(
    month = as_date(month),
    year = floor_date(month, unit = "year"),
    zip_code = as.numeric(zip_code)) |> 
  select(-county_name)

zip_codes_df = 
  read_csv("./zillow_data/Zip Codes.csv") |> 
  janitor::clean_names() |> 
  filter(
    !(zip_code == 10463 & county == "New York"),
    !(zip_code == 11201 & county == "New York"))  |> 
  mutate(
    borough = case_match(
      county,
      "Bronx" ~ "Bronx",
      "Kings" ~ "Brooklyn",
      "New York" ~ "Manhattan",
      "Queens" ~ "Queens",
      "Richmond" ~ "Staten Island"
    )
  ) |> 
  select(zip_code, borough, neighborhood)

nyc_price_df = 
  left_join(nyc_zillow_df, zip_codes_df, by = "zip_code")
```

```{r}
Q2=
  nyc_price_df |> 
  filter(month >= as.Date("2015-01-01"), month <= as.Date("2024-08-31")) |> 
  group_by(zip_code) |> 
  summarize(n_months = sum(!is.na(price))) |> 
  summarize(
    all = sum(n_months == 116),
    less10 = sum(n_months < 10)
  )

```
There are `r Q2$all` codes are observed in all 116 months.There are `r Q2$less10` codes are observed fewer than 10 times.

```{r}
nyc_price_df |>
  group_by(borough, year) |>
  summarize(avg_rental_price = mean(price, na.rm = TRUE)) |>
  pivot_wider(
    names_from = year,
    values_from = avg_rental_price
  ) |>
  arrange(borough) |>
  knitr::kable()
```
According to the table, the average rental price in each borough increase every year.

```{r}
smooth=
nyc_price_df |>
  ggplot(aes(x = year, y = price, color = borough)) +
  geom_smooth()+
   labs(
    title = "NYC Rental Prices (2015–2024)",
    x = "Year",
    y = "Average Rental Price",
    caption = "Data from zillow",
    color = "Borough"
  ) +
  theme(legend.position="bottom")

```
The polt shows that there is a increase trend in rental prices in all boroughs from 2015 to 2024 in NYC. Manhattan have the hightest price and Bronx have the lowest price over the year.

```{r}
violin=
nyc_price_df |>
  filter(year == "2023-01-01") |>
  group_by(borough, zip_code) |>
  summarize(mean_price = mean(price, na.rm = TRUE)) |>
  ggplot(aes(x = borough, y = mean_price)) +
  geom_violin(aes(fill = borough), alpha=.5) +
  stat_summary(fun="mean",color="blue")
  labs(
    title = "Average Rental Price within ZIP Code in 2023",
    x = "Borough",
    y = "Average Rental Price"
  ) +
 theme(legend.position="bottom")

```
Like the last polt, Manhattan have the highest average rents in NYC, however, it also have the widest range.Staten Island have the narrowest range in NYC.

```{r}
library(patchwork)
smooth/violin+
  labs(
    title = "Average Rental Price within ZIP Code in 2023",
    x = "Borough",
    y = "Average Rental Price"
  )
ggsave("p8105_hw1p2_Combine.pdf")
```

##Problem 3

